# (PART) Acting on common data structures {-} 

# Data frames and tibbles

## Basic operations and related stuff

**Task:** Select a column and return it as vector (not data.frame).

```{r, eval = FALSE}
iris %>% .$Species

# or

iris %>% dplyr::pull(Species)
#   ____________________________________________________________________________
iris[["Species"]]
```

**Task:** Select columns from (and return) data.frame.

```{r, eval = FALSE}
iris %>% select(Sepal.Width, Species)
#   ____________________________________________________________________________
iris[, c("Sepal.Width", "Species"), drop = FALSE]
```

**Task:** Filter/subset a data.frame.

```{r, eval = FALSE}
iris %>% filter(Species == "versicolor" & Sepal.Width < 3)
#   ____________________________________________________________________________
iris[iris[, "Species"] == "versicolor" & iris[, "Sepal.Width"] < 3]
```

**Task:** Mutate/transform/calculate new columns of a data.frame.

```{r, eval = FALSE}
iris %>% mutate(a = Sepal.Length, b = 2 * a)
#   ____________________________________________________________________________
iris[["a"]] <- iris[["Sepal.Length"]]
iris[["b"]] <- 2 * iris[["a"]]
```

**Task:** Arrange/order a data.frame.

```{r, eval = FALSE}
dplyr::arrange(iris, dplyr::desc(Sepal.Length))
#   ____________________________________________________________________________
iris[order(iris["Sepal.Length", ], decreasing = TRUE), ]
```

**Task:** Rename colums

```{r, eval = FALSE}
iris %>%
  dplyr::rename(a = Species,
                b = Sepal.Length)
# or purrr::set_names()
# or dplyr::select()
#   ____________________________________________________________________________
# Use names() or setNames()
```

**Task:** Group and summarise a data.frame (#annonymous function).

```{r}
iris %>% dplyr::group_by(Species) %>% 
  dplyr::summarise(n_rows = n(),
                   dists = n_distinct(Sepal.Width, Sepal.Length),
                   blub = (function(x) mean(x))(Sepal.Width),
                   m = mean(Sepal.Length))
#   ____________________________________________________________________________
```

## Reshaping wide and long data

**Task:** Gather/melt/unpivot/fold from wide into long format (two columns).

```{r}
data.frame(abc = sample(letters[1:3], 6, replace = TRUE), 
           r1 = rnorm(6),
           r2 = rnorm(6),
           stringsAsFactors = FALSE) %>% 
  tidyr::gather(key = r, value = random, r1, r2) %>% 
  head(4)
#   ____________________________________________________________________________
```

**Task:** Spread/cast/pivot/unfold from long into wide format (two columns).

```{r}
data.frame(abc = rep(letters[1:4], times = 2),
           r = rep(c("r1", "r2"), each = 4),
           random = rnorm(8),
           stringsAsFactors = FALSE) %>% 
  tidyr::spread(key = abc, value = random) %>% 
  head(1)
#   ____________________________________________________________________________
```

## Typical mutate statements

### Split and paste

**Task:** Split/separate a column into one or more new columns.

```{r}
iris %>% dplyr::transmute(Sepal.Length = as.character(Sepal.Length)) %>%
  tidyr::separate(col = Sepal.Length, 
                  into = c("Sep1", "Sep2"),
                  sep = "\\.",
                  remove = FALSE) %>% 
  head(1)
```

**Task:** Combine/unite/paste one or more columns into one.
```{r}
iris[1,] %>% tidyr::unite(new_col, Sepal.Width, Species, sep = "", remove = FALSE)
iris[1,] %>% dplyr::mutate(new_col = stringr::str_c(Sepal.Width, Species))
#   ____________________________________________________________________________
```

### Conditionals

**Task:** Mutate columns depending/conditionally on other colums.

```{r}
# use dplyr::case_when() or dplyr::if_else()
#   ____________________________________________________________________________
# use ifelse
```

### IDs

**Task** Add an id column.

```{r}
iris[1:2, ] %>% 
  dplyr::mutate(id = seq_len(n()))
#   ____________________________________________________________________________
df <- data.frame(a = 1:2)
df[["id"]] <- seq_len(nrow(df))
```

**Task** Add a unique identifier regarding some columns (not in order).

```{r, eval = FALSE}
iris %>% 
  dplyr::mutate(id = as.integer(factor(stringr::str_c(Species, Sepal.Width))))
#   ____________________________________________________________________________
iris[["id"]] <- as.integer(factor(paste0(iris[["Species"]], iris[["Sepal.Width"]])))
```

**Task** Add a unique identifier regarding all (or some) columns, in order off appearing unique rows.

```{r, eval = FALSE}
iris %>% dplyr::mutate(id = iris %>% 
                         purrr::pmap_chr(stringr::str_c, sep = "\t") %>% 
                         factor %>% 
                         forcats::fct_inorder(ordered = TRUE) %>% 
                         as.integer)
#   ____________________________________________________________________________
```

### Other stuff

**Task:** Use some helpers to select columns based on naming patterns

```{r}
iris[1,] %>%
  dplyr::select(dplyr::starts_with("S",
                                   ignore.case = TRUE,
                                   vars = dplyr::current_vars()))
iris[1,] %>% dplyr::select(dplyr::ends_with("h"))
iris[1,] %>% dplyr::select(dplyr::contains("."))
iris[1,] %>% dplyr::select(dplyr::matches("."))
purrr::set_names(iris[1,], paste0("a", 1:5)) %>%
  dplyr::select(dplyr::num_range(prefix = "a", range = 2:4, width = 1))
iris[1, ] %>% dplyr::select(dplyr::one_of("Sepal.Width", "Species"))
#   ____________________________________________________________________________
```

**Task:** Summarise grouped data by a statistic that returns more than one value.

**Task:** Add summary data directly to the summarised data frame.


## Formatting
**Task:** Change column names to lower case.

```{r}
stringr::str_to_lower(names(iris))
#   ____________________________________________________________________________
tolower(names(iris))
```

**Task:** Change column names to snake_case.

```{r}
library(magrittr)
c("Var 1", "Var-2", "Var.3", "Var4") %>%
  stringr::str_replace_all("\\s|-|\\.", "_") %>% 
  stringr::str_to_lower()
```

**Task:** Change column names from camelCase to snake_case.

```{r}
camelCases <- c("smallCamelCase", "BigCamelCase", "mixed_Case", "snake_case")

to_snake_case <- function(camelcases){
  # get to know, if a string starts with a small letter
  small_start <- !is.na(stringr::str_extract(camelCases, "^[a-z]"))
  # get all capital letters from a string
  capitals <- stringr::str_extract_all(camelCases, "[A-Z]")
  # Setting an underscore before capital and first letters
  starts <- purrr::pmap(list(camelCases,
                             small_start,
                             capitals),
                        function(x,y,z)
                          if (length(z) == 0) {"_"} else {
                            c("_", paste0("_", z))
                            }
                        )
  # split the strings by their capital letters.
  rests <- stringr::str_split(camelCases, "[A-Z]")
  # setting all peaces together:
  # - pasting first and capital letters with the rest of the string
  # - applying tolower, remove more than one "_" and starting "_"
  corrected <- purrr::map2_chr(starts, rests, stringr::str_c, collapse = "") %>% 
    purrr::map_chr(stringr::str_to_lower) %>% 
    purrr::map_chr(~ stringr::str_replace_all(.x, "_+", "_")) %>% 
    purrr::map_chr(~ stringr::str_replace(.x, "^_", ""))
  corrected
}

to_snake_case(camelCases)
```

**Task::** Change column names from snake_case to camelCase.

**Task:** Mutate/Change/reorder columns depending on the row number (for example the last row).

```{r}
iris %>% head(1) %>% dplyr::select(Species, dplyr::everything())
#   ____________________________________________________________________________
setNames(iris[1, ], c("Species", setdiff(names(iris), "Species")))
```


**Task** Work on more than two tables (`Reduce()`)

**Task** Convert rownames of a data frame into first column.


## Joins

**Task:** Join on two columns.

## Tibbles

## Resources

* https://www.r-bloggers.com/lesser-known-dplyr-tricks/
