# (PART) Acting on common data structures {-} 

# Data frames and tibbles


## Formatting
**Task:** Change column names to lower case.

**BASE:**
```{r}
tolower(names(iris))
```

**Task:** Change column names from camelCase or CamelCase to snake_case and reversed.

## Basic operations and related stuff

**Task:** Select columns from a data.frame.

**TDVS:** `iris %>% select(Sepal.Width, Species)`

**BASE:** `iris[, c("Sepal.Width", "Species")]`

**Task:** Filter/Subset a data.frame.

**TDVS:** `iris %>% filter(Species == "versicolor" & Sepal.Width < 3)`

**BASE:** `iris[iris[, "Species"] == "versicolor" & iris[, "Sepal.Width"] < 3]`

**Task:** Mutate/transform/calculate new columns of a data.frame.

### Arrange/order a data.frame.
#### TDVS
```{r}
dplyr::arrange(iris, dplyr::desc(Sepal.Length))
```

#### BASE
```{r}
iris[order(iris["Sepal.Length", ], decreasing = TRUE), ]
```


**Task:** Split a column into one or more new columns.

**Task:** Combine/unite one or more columns into one.

**Task:** Group and summarise a data.frame.

## Common use cases

**Task:** Mutate/Change columns depending on the row number (for example the last row).

**TDVS:** `mutate(if_else(row_number() == n(), 1, 0))`

**Task** Work on more than two tables (`Reduce()`)

**Task** Convert rownames of a data frame into first column.

## Wide and Long data

**Task:** Gather/Melt from wide into long format (two columns).

**Task:** Reshape/Spread from long into wide format (two columns).

## Joins

**Task:** Join on two columns.

## Measures

**Task:** Summarise grouped data by a statistic that returns more than one value.

## Tibbles

## Resources
