# Factors

**Task:** Create a factor.

```{r, eval = TRUE}
#   ____________________________________________________________________________
factor(letters)
```

**Task:** Check if an atomic vector is a factor.

```{r, eval = TRUE}
#   ____________________________________________________________________________
is.factor(factor(letters))
```

**Task:** Create an ordered factor.

```{r, eval = TRUE}
#   ____________________________________________________________________________
factor(letters, ordered = TRUE)
```

**Task:** Check if a factor is ordered.

```{r, eval = TRUE}
#   ____________________________________________________________________________
is.ordered(factor(letters))
```

**Task:** Add a factor level.

```{r, eval = TRUE}
#   ____________________________________________________________________________
a <- factor(letters[1:5])
levels(a) <- c(levels(a), "new_level")
a
```

**Task:** Add a value (that is not a level of the factor).

```{r, eval = TRUE}
#   ____________________________________________________________________________
a <- factor(letters[1:5])
a <- factor(c(as.character(a), "new_level"))
a
```

**Task:** Combine two factors.

```{r, eval = TRUE}
#   ____________________________________________________________________________
a <- factor(letters[1:5])
b <- factor(letters[3:7])
factor(c(as.character(a), as.character(b)))
```

**Task:** Sort the levels of a factor by another variable.

**Task:** Reorder the levels of a factor by another variable.

**Task:** Reverse the order of factor levels.

**Task:** Change specific levels of a factor by name, while preserving the order.

```{r, eval = TRUE}
a <- factor(letters[1:5])
forcats::fct_recode(a, g= "a", h = "b")
# Note that the folowing will not work, since f is also the argument name of the 
# data: forcats::fct_recode(a, f= "a", g = "b")
```

**Task:** Delete a factor level.

```{r, eval = TRUE}
a <- factor(letters[1:5])
forcats::fct_recode(a, NULL= "a", h = "b")
```

**Task:** Move specific factor levels in front of all others.

**Task:** Count factors/strings.

```{r, eval = TRUE}
a <- factor(rep(letters[1:5], 5:1))
forcats::fct_count(a)
```

**Task:** Join/bind/combine/compose/unite rare/common factor levels into one.

```{r, eval = TRUE}
a <- rep(letters[1:5], 5:1)
forcats::fct_lump(a, n = 2)
forcats::fct_lump(a, n = -2)
```

## Resources

* [forcats 0.1 on RStudio blog](https://blog.rstudio.org/2016/08/31/forcats-0-1-0/)
