```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
# RStudio project management

## Project types

## Project options

## Version control

## Packrat and package dependencies

Resources: 

* [Packrat Webinar by RStudio](https://www.rstudio.com/resources/webinars/managing-package-dependencies-in-r-with-packrat/)
* [Packrat Homepage](https://rstudio.github.io/packrat/)
    
Most important commands:

* `packrat::init()`, initializes an isolated package library for the current project.
* `packrat::snapshot()` saves all changes within the packrat project library since the last snapshot.
* `packrat::restore()` deletes all changes within the packrat project library since the last snapshot.

## Addins

1. **Task:** Structure/separate code into blocks by different lines of lentgh 80 (by default).

    __<span style="color:grey">General:</span>__ Use `devtools::install_github("lorenzwalthert/strcode")` within your project (to install the `strcode` package from [here](https://github.com/lorenzwalthert/strcode)). That will enable the following:
    
    ```{r, eval = FALSE}
    strcode::insert_l1_break() # or strg + c + 1 (shortcut)
    #   ____________________________________________________________________________
    strcode::insert_l2_break() # or strg + c + 1
    ##  ............................................................................
    strcode::insert_l3_break() # or strg + c + 1
    ### .. . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . .
    ```
