```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
# RStudio project management

## Project types

Empty project, package or shiny web applications

## Project options



## Version control

* copy the link from a remote repository, when creating a new package
* git add .
* git add -m "some commit message"
* push to remote

## Packrat and package dependencies

Resources: 

* [Packrat Webinar by RStudio](https://www.rstudio.com/resources/webinars/managing-package-dependencies-in-r-with-packrat/)
* [Packrat Homepage](https://rstudio.github.io/packrat/)
    
Most important commands:

* `packrat::init()`, initializes an isolated package library for the current project.
* `packrat::snapshot()` saves all changes within the packrat project library since the last snapshot.
* `packrat::restore()` deletes all changes within the packrat project library since the last snapshot.
* `packrat::bundle()` creates a tarball of the project, the packages and its dependencies.
* `packrat::unbundle(bundle = "path//to//the//bundle", where = ".")` unbundles (restores) a project within a subdirectory of another (conveniently empty) project.
* `packrat::off()` lets you temporarily step out of "packrat mode", which means, that you will now use your global user library instead of the packrat project library.
* `packrat::on()` lets you go back into "packrat mode".

Collaboration:

* Initialise a packrat project on github, let everyone clone it and let them then call `packrat::restore()`.

Updated R Version:

* When you unbundle a project, which was created under an older (or newer) R version, then installed, packrat will warn you and install the older (or newer) R version when calling `packrat::snapshot()`. This is one among other reasons, why it is easier to work with packrat, while having access to the internet.

## Addins

1. **Task:** Structure/separate code into blocks by different lines of lentgh 80 (by default).

    __<span style="color:grey">General:</span>__ Use `devtools::install_github("lorenzwalthert/strcode")` within your project (to install the `strcode` package from [here](https://github.com/lorenzwalthert/strcode)). That will enable the following:
    
    ```{r, eval = FALSE}
    strcode::insert_l1_break() # or strg + c + 1 (shortcut)
    #   ____________________________________________________________________________
    strcode::insert_l2_break() # or strg + c + 1
    ##  ............................................................................
    strcode::insert_l3_break() # or strg + c + 1
    ### .. . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . .
    ```
