```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)

Sys.setenv(LANGUAGE = "en") 
```
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)

Sys.setenv(LANGUAGE = "en") 
```
# Spatial data

## IO

1. **How to load/read and save/write spatial data?**

```{r}
library(sf)
## Linking to GEOS 3.5.1, GDAL 2.1.3, proj.4 4.9.2, lwgeom 2.3.2 r15302
nc <- st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)
# st_write(nc, "nc.shp")
# see st_read_db() and st_write_db() for spatial databases
# look carefully through all of the arguments (especially layers,
# stringsAsFactors, geometry_column, type, quiet)
# in case of specific problems see:
# https://edzer.github.io/sfr/articles/sf2.html
#   ____________________________________________________________________________
```

## Objects

2. **What are the classes of sf-objects?**

```{r}
# the data.frame
class(nc)

# the geometry list-column
class(st_geometry(nc))

# the individual feature geometries
lapply(st_geometry(dplyr::slice(nc, 1:2)), class)
```

3. **Which methods are avaliable for sf-objects**?

```{r}
methods(class = "sf")

methods(class = 'sfc')

methods(class = 'sfg')
```

4. **What sf-geometry types are available, how do you initialize them and what types are the internally?**

```{r, eval = FALSE}
# internal types:
# - point (atomic)
# - linestring (matrix)
# - everything else (list)

## common geometries
# point
st_point(c(1,2))
# linestring
st_linestring(rbind(c(0.2,3), c(0.2,4), c(1,4.8), c(2,4.8)))
# polygon
st_polygon(list(
  rbind(c(0,0), c(1,0), c(3,2), c(2,4), c(1,4), c(0,0)),
  rbind(c(1,1), c(1,2), c(2,2), c(1,1))))
# multipoint
st_multipoint(rbind(c(3.2,4), c(3,4.6), c(3.8,4.4), c(3.5,3.8), c(3.4,3.6)))
# multilinestring
st_multilinestring(list(rbind(c(0.2,3), c(0.2,4), c(1,4.8), c(2,4.8)),
                        rbind(c(0,4.4), c(0.6,5))))
# multipolygon
st_multipolygon(list(
  list(rbind(c(0,0), c(1,0), c(3,2), c(2,4), c(1,4), c(0,0)),
       rbind(c(1,1), c(1,2), c(2,2), c(1,1))),
  list(rbind(c(3,0), c(4,0), c(4,1), c(3,1), c(3,0)),
       rbind(c(3.3,0.3), c(3.8,0.3), c(3.8,0.8), c(3.3,0.8), c(3.3,0.3))[5:1,]),
  list(rbind(c(3,3), c(4,2), c(4,3), c(3,3)))))

# geometrycollection
st_sfc(st_geometrycollection(list(st_point(1:2))), 
       st_geometrycollection(list(st_linestring(matrix(1:4,2)))))
# geometry
st_sfc(st_point(1:2), st_linestring(matrix(1:4,2))) # like the latter, but as a "superclass"

## rare geometries
# circularstring
# compoundcurve
# curvepolygon
# multicurve
# multisurface
# curve
# surface
# polyhedral surface
# tin
# triangle
```

5. **What are the dimensions of those objects?**

```{r}
x = st_sfc(
  st_point(0:1), 
  st_linestring(rbind(c(0,0),c(1,1))), 
  st_polygon(list(rbind(c(0,0),c(1,0),c(0,1),c(0,0)))),
  st_multipoint(),
  st_linestring(),
  st_geometrycollection())
st_dimension(x, NA_if_empty = FALSE)
```

5. **How to initialize these objects from well known text?**

```{r}
st_as_sfc("POINT(0 1)")
```

5. **How to initialize an sf-object and bring it into coordinate reference systems (crs)?**

```{r}
# attributes of sfc obejects to store a crs: epsg, proj4string
# theses must be the same for all elements of an sfc-object, but can also be NA.
# -proj4string is a generic, string-based description of crs
# - epsg integer values refer to particular well known crs, that might also
# improve over time

st_sf(points = st_sfc(st_point(c(0, 0)),
                      st_point(c(0, 1)),
                      crs = 4326))
# also use:
# use st_crs() and st_transform()
```

6. **How to coerce data.frames to sf?**

```{r}
# use st_as_sf
```

6. **How to create and identify empty geometries?**

```{r}
st_geometrycollection()
length(st_geometrycollection())
```

7. **How to identify unconnected parts of geometries?**

```{r}
which(sapply(st_geometry(nc), length) > 1)
```

8. **How to switch between well-known text and well-known binary encoding?**

```{r}
x <- st_linestring(matrix(10:1,5))
st_as_text(x)
st_as_binary(x)
```

9. **How to control geometries for simplicity and validity?**

```{r}
st_is_valid(st_as_sfc("POLYGON((0 0, 0 10, 10 0, 10 10, 0 0))"))
st_is_simple(st_as_sfc("POLYGON((0 0, 0 10, 10 0, 10 10, 0 0))"))
plot(st_as_sfc("POLYGON((0 0, 0 10, 10 0, 10 10, 0 0))"))
```

9. **How to create a grid manually?**

```{r}
library(dplyr)
lat <- seq(from = 1, to = 50, length.out = 100)
long <- seq(from = 1, to = 50, length.out = 100)
grid <- tidyr::expand(tibble(lat = lat, long = long), lat, long)
grid <- grid %>% 
  mutate(points = purrr::map2(lat, long,
                              ~st_geometrycollection(list(st_point(c(.x,.y))))))
grid <- grid %>% mutate(points = st_sfc(points))
grid <- st_sf(grid, stringsAsFactors = FALSE,
              sf_column_name = "points")
plot(grid %>% select(points))
```

**How to set aggregations?**

```{r}
# while initializing
nc <- st_read(system.file("shape/nc.shp", package="sf"),
    agr = c(AREA = "aggregate", PERIMETER = "aggregate", CNTY_ = "identity",
        CNTY_ID = "identity", NAME = "identity", FIPS = "identity", FIPSNO = "identity",
        CRESS_ID = "identity", BIR74 = "aggregate", SID74 = "aggregate", NWBIR74 = "aggregate",
        BIR79 = "aggregate", SID79 = "aggregate", NWBIR79 = "aggregate"))
```

## Arithmetics

10. **How to calculate with simple features?**

```{r}
st_polygon(list(rbind(c(0,0),c(1,0),c(1,1),c(0,1),c(0,0)))) - 0.5
st_point(c(0,0)) + 1
```

10. **How to calculate distances?**

```{r}
x = st_transform(nc, 32119)
st_distance(x[c(1,4,22),], x[c(1, 33,55,56),], dist_fun = geosphere::distGeo)
line = st_sfc(st_linestring(rbind(c(30,30), c(40,40))), crs = 4326)
st_length(line)
```

## Plots

10. **How does the plot method and its `add` argument work?**

```{r}
plot(nc[1,1], col = 'grey')
```

```{r}
plot(nc[1,1], col = 'grey')
plot(nc[1], add = TRUE)

plot(nc[1])
plot(nc[1,1], col = 'grey', add = TRUE)
```

11. **Some differenct plot options**

```{r}
plot(nc[1])
plot(st_convex_hull(nc[1]))
plot(nc[1], border = grey(.5))
```

## Resources

* [sf-page](https://edzer.github.io/sfr/index.html)
* [tidy spatial data in R with ... (blogpost)](http://strimas.com/r/tidy-sf/)
* [sf-issue: dplyr compatibility](https://github.com/edzer/sfr/issues/42)
